Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-03-24 2_SSL_EES/article/article_main.tex
vaaarious fixes

%%%TITLE, AUTHORS AND ABSTRACT%%%
\twocolumn[
  \begin{@twocolumnfalse}
{\includegraphics[height=30pt]{auxiliary/rsc_template_head_foot/journal_name}\hfill\raisebox{0pt}[0pt][0pt]{\includegraphics[height=55pt]{auxiliary/rsc_template_head_foot/RSC_LOGO_CMYK}}\\[1ex]
\includegraphics[width=18.5cm]{auxiliary/rsc_template_head_foot/header_bar}}\par
\vspace{1em}
\sffamily
\begin{tabular}{m{4.5cm} p{13.5cm} }

@@

 & \noindent\large{Michael Weinold \textit{$^{a b \ddag}$}, Sergey Kolesnikov,\textit{$^b$} and Laura Diaz Anadon\textit{$^{b}$}} \\

@@

\footnotetext{\textit{$^{a}$~ETH Zurich, Zurich, Switzerland. E-mail: michael.weinold@alumni.ethz.ch}}
\footnotetext{\textit{$^{b}$~Cambridge Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, United Kingdom. }}

@@

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies}\cite{Anadon2016}. Among these drivers, the role of external  knowledge transfer and technology spillovers in research and development of energy technologies is an understudied area  \cite{Stephan2021}. While the exact definition of spillovers in the literature depends on the context and the research question \cite{Liu2003}\cite{Nemet2012}, we follow Stephan et al. \cite{Stephan2021} and consider knowledge to be external - thus a spillover -  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network outside a technology may be an important factor in understanding \cite{Pichler2020} and shaping \cite{Clark2016}\cite{Stephan2021}\cite{Sun2021}\cite{kolesnikov_framework_2022} the future evolution of technologies.

@@

The historical development XXXX development into powerful lghting devices from very dim novelty XXXX of light-emitting diodes received a considerable amount of attention following the recognition of pioneering work on blue LED by Japanese researchers Shuji Nakamura, Isamu Akasaki and Hiroshi Amano with the 2014 Nobel Prize in Physics \cite{Akasaki2015}\cite{Nakamura2015}. The sources and disaggregated contributions of subsequent improvements in LED technology, however, have not been documented in the literature in a systematic fashion, though there were notable publications reporting on the progress in the design of devices \cite{Shchekin2006}\cite{krames2007led}\cite{laubsch2009high}\cite{hahn2014development}, or the technological improvements  underlying the improvement in overall device efficiency for best performers in 2009 \cite{tsao2010solid} and 2016 \cite{pattison2017solid}.

@@

To supplement our data collection efforts, verify our findings and identify additional spillovers, we conducted a series of elite semi-structured interviews with eleven eminent experts from academia, industry and the public research sector. Experts were selected based on their engagement in  different sub-fields of LED research and manufacturing, as well as the recommendations from other interviewed experts, in essence expanding the list of experts that emerged from the initial literature review. All interviews were conducted between November 2019 and April 2022 by means of video conferencing and lasted for about one hour. A summary of the background of interviewed experts is provided in Table SI 3 in Section 5 of the Supplementary Information.

@@

Figure \ref{fgr:breakthroughs_efficiency} shows the overall magnitude of contributions of identified LED innovations and technology spillovers to improvements in LED efficiency over time across different sub-efficiencies. The list of technology spillovers identified in our study is provided in Table \ref{tab:spillovers} and discussed in Section \ref{subsec:spillovers}. The complete list of all LED innovations is provided in Table SI 2 in Section 3 of the Supplementary Information. Through the index decomposition analysis described in Section \ref{subsec:metrics}, we find that out of the overall LED efficiency increase of 32.9\% from 5.8\% to 38.8\% between 2003 and 2020, at least 2.8\% can be attributed specifically to technology spillovers identified in this study, corresponding to 8.5\% of the total LED efficiency improvements between 2003 and 2020.

@@

\begin{table*}[h!]
    \small
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3 \\ \hline
    \end{tabularx}
    \caption{LED down-conversion materials (phosphors and quantum dots) related to improvements in consumer experience metrics identified in this study. The \textit{year} column represents the earliest reported application of invention in white LEDs. These differ from the years used in Figure \ref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for representative LED products that relied on those materials. The \textit{SI Sec.} column refers to the corresponding section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Data points show the color temperature and colour rendering index performance of the earliest identified representative white LED products with published spectral data that used phosphors listed in Table \ref{tab:phosphors}, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI 9 in section 4 of the Supplementary Information. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. The desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

>>>>

%%%TITLE, AUTHORS AND ABSTRACT%%%
\twocolumn[
  \begin{@twocolumnfalse}
{\includegraphics[height=30pt]{auxiliary/rsc_template_head_foot/journal_name}\hfill\raisebox{0pt}[0pt][0pt]{\includegraphics[height=55pt]{auxiliary/rsc_template_head_foot/RSC_LOGO_CMYK}}\\[1ex]
\includegraphics[width=18.5cm]{auxiliary/rsc_template_head_foot/header_bar}}\par
\vspace{1em}3
\sffamily
\begin{tabular}{m{4.5cm} p{13.5cm} }

@@

 & \noindent\large{Michael Weinold \textit{$^{a, b \ddag}$}, Sergey Kolesnikov,\textit{$^b$} and Laura Diaz Anadon\textit{$^{b, c}$}} \\

@@

\footnotetext{\textit{$^{a}$~ETH Zurich, Zurich, Switzerland. E-mail: michael.weinold@alumni.ethz.ch}}
\footnotetext{\textit{$^{b}$~Cambridge Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, United Kingdom.}}
\footnotetext{\textit{$^{c}$~Belfer Center for Science and International Affairs, Harvard University, Cambridge, MA, United States.}}

@@

The stages and drivers of the innovation life-cycle for a particular technology play out within a broader innovation system \cite{grubler2012policies}\cite{Anadon2016}. Among these drivers, the role of external  knowledge transfer and technology spillovers in research and development of energy technologies is an understudied area  \cite{Stephan2021}. While the exact definition of spillovers in the literature depends on the context and the research question \cite{Liu2003}\cite{Nemet2012}, we follow Stephan et al. \cite{Stephan2021} and consider knowledge to be external - thus a spillover -  if it has been developed for application in other technologies, sectors, or scientific disciplines. There is emerging evidence that understanding spillovers and the knowledge network outside a technology may be an important factor in understanding \cite{Pichler2020} and shaping \cite{Clark2016}\cite{Stephan2021}\cite{Sun2021}\cite{kolesnikov2022technology} the future evolution of technologies.

@@

The historical development development of light-emitting diodes from a novelty semiconductor experiment into powerful lighting devices has received a considerable amount of attention following the recognition of pioneering work on blue LED by Japanese researchers Shuji Nakamura, Isamu Akasaki and Hiroshi Amano with the 2014 Nobel Prize in Physics \cite{Akasaki2015}\cite{Nakamura2015}. The sources and disaggregated contributions of subsequent improvements in LED technology, however, have not been documented in the literature in a systematic fashion, though there were notable publications reporting on the progress in the design of devices \cite{Shchekin2006}\cite{krames2007led}\cite{laubsch2009high}\cite{hahn2014development}, or the technological improvements  underlying the improvement in overall device efficiency for best performers in 2009 \cite{tsao2010solid} and 2016 \cite{pattison2017solid}.

@@

To supplement our data collection efforts, verify our findings and identify additional spillovers, we conducted a series of elite semi-structured interviews with eleven eminent experts from academia, industry and the public research sector. Experts were selected based on their engagement in  different sub-fields of LED research and manufacturing, as well as the recommendations from other interviewed experts, in essence expanding the list of experts that emerged from the initial literature review. All interviews were conducted between November 2019 and April 2022 by means of video conferencing and lasted for about one hour. A summary of the background of interviewed experts is provided in Table SI 3 in Section 6 of the Supplementary Information.

@@

Figure \ref{fgr:breakthroughs_efficiency} shows the overall magnitude of contributions of identified LED innovations and technology spillovers to improvements in LED efficiency over time across different sub-efficiencies. The list of technology spillovers identified in our study is provided in Table \ref{tab:spillovers} and discussed in Section \ref{subsec:spillovers}. The complete list of all LED innovations is provided in Table SI 2 in Section 4 of the Supplementary Information. Through the index decomposition analysis described in Section \ref{subsec:metrics}, we find that out of the overall LED efficiency increase of 32.9\% from 5.8\% to 38.8\% between 2003 and 2020, at least 2.8\% can be attributed specifically to technology spillovers identified in this study, corresponding to 8.5\% of the total LED efficiency improvements between 2003 and 2020.

@@

\begin{table*}[h!]
    \small
    \centering
    \begin{tabularx}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{SI Sec.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 5.1.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 5.1.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 5.1.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 5.1.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 5.1.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 5.1.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 5.1.2 \\ \hline
    \end{tabularx}
    \caption{LED down-conversion materials (phosphors and quantum dots) related to improvements in consumer experience metrics identified in this study. The \textit{year} column represents the earliest reported application of invention in white LEDs. These differ from the years used in Figure \ref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for representative LED products that relied on those materials. The \textit{SI Sec.} column refers to the corresponding section in the Supplementary Information that contains references and a detailed description of the history of corresponding invention. Abbreviations: Desig. - Designation, Sec. - Section.}
    \label{tab:phosphors}
\end{table*}

\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/breakthroughs_consumer-experience.pdf}
 \caption{Historical improvements in consumer experience metrics of phosphor-converted white LEDs. Data points show the color temperature and colour rendering index performance of the earliest identified representative white LED products with published spectral data that used phosphors listed in Table \ref{tab:phosphors}, each indicated by the phosphor label and publication year. Corresponding spectral data used to calculate colour temperature values is provided in Figure SI 11 in section 5 of the Supplementary Information. Three data points representing compact fluorescent lamps (CFLs)\cite{cie_reference}, shown as black triangles, are provided for comparison. Horizontal lines represent typical colour temperatures of “traditional” light sources, shown for reference. The desirable range of colour temperatures for home illumination, indicated by a vertical red arrow, lies between two horizontal orange lines representing typical incandescent light and warm daylight colour temperatures. Horizontal red arrow indicates desirable higher values of colour rendering index (CRI). LED products based on the following phosphor innovations from the following LED manufacturers are represented : YAG - Nichia, 1998 \cite{bando1998development}; YGAG - Lumileds, 2002 \cite{Mueller2002}; 258 - Lumileds, 2005 \cite{MuellerMach2005}; SLA - Lumileds, 2014 \cite{Pust2014}; PFS - GE, 2015 \cite{Murphy2015}; QD - Lumileds, 2017 \cite{lumileds2016qd}, Osram, 2019 \cite{osram2019qd}; SALON - Osram, 2019 \cite{Hoerder2019}.}
 \label{fgr:consumer_experience}
\end{figure*}

### 2023-03-24 2_SSL_EES/article/article_supplementary.tex
vaaarious fixes

The manufacturing process of semiconductor devices can be categorized by the level at which manufacturing process steps are implemented, i.e., either at the wafer level or at the individual chip/package level. The total manufacturing cost per die is thus the sum of the total costs of all wafer processing steps and all die packaging steps.
\begin{myequation}
\label{eqn:cost_sum}
    C \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = P_S + C_w + C_p
\end{myequation}
where
\begin{align*}
    P_S &\dots \text{sapphire substrate price per die} \\
    C_w &\dots \text{wafer processing cost per die} \\
    C_p &\dots \text{die processing cost}
\end{align*}
The total wafer processing cost and total die packaging costs are in turn the sum of all associated process steps.
\begin{myequation}
        C_w = \sum_i C_i
\end{myequation}
\begin{myequation}
	C_p = \sum_j C_j
\end{myequation}
The cost of a single process step $C_i$ can now be written as
\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x c_x \bigg\}
\end{myequation}
where the index $i$ runs over all wafer processing steps, the index $j$ runs over all die processing steps and the index $x$ run over all materials.
\begin{align*}
        D &\dots \text{number of functional (i.e., successfully tested) die per wafer}\\
        y &\dots \text{process step yield} \\
        u &\dots \text{equipment utilization (relative to theoretical equipment capacity)} \\
        p &\dots \text{power consumption} \\
        c_e &\dots \text{hourly electricity cost} \\
        c_m &\dots \text{hourly maintenance cost} \\
        c_d &\dots \text{hourly depreciation cost}\\
        c_l &\dots \text{hourly labour cost} \\
        c_o &\dots \text{hourly overhead cost} \\
        t_i &\dots \text{time per run} \\
        w &\dots \text{wafers per run} \\
        A &\dots \text{wafer area} \\
        v_x &\dots \text{volume of material $x$ per wafer} \\
        c_x &\dots \text{cost of material $x$ per volume}\\
\end{align*}
The number of die per wafer $D$ depends on the total usable wafer area. The usable area depends on the wafer diameter, the cutting street width between the chips and the exclusion zone at the rim of the wafer.
\begin{myequation}
	A_{\text{usable}}=A_{\text{wafer}}-A_{\text{cut}}-A_{\text{exclusion}}
\end{myequation}
Determining the usable wafer area as a function of these three parameters requires a numerical solution. However, following discussions in the literature \cite{de2005investigation}, we approximate the number of functional die per wafer\footnote{often \textit{"good die per wafer"} in the literature} as
\begin{myequation}
\label{eqn:dpw}
	DPW = \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2}
\end{myequation}
where
\begin{align*}
    d &\dots \text{wafer diameter} \\
    e &\dots \text{wafer edge exclusion zone width} \\
    a &\dots \text{die area} \\
    s &\dots \text{cutting street width} \\
\end{align*}

@@

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW) over time. \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category
\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x c_x \bigg\}
\end{split}
\end{myequation}
and the cost of a manufacturing step $C_j$ in the packaging category
\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x c_x \bigg\}
\end{myequation}
where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}
Note that in keeping with the categorization introduced by the United States Department of Energy (cf. eg. \cite{doe_ssl_rnd_2019}), certain steps from these two categories are reported separately. In the wafer processing category, the epitaxy step is reported separately due to its complexity and the large share of cost carried. In the wafer processing category, the phosphor step is reported separately.

@@

@@

            ongoing & Epitaxy improvements \newline and better doping & polarization and \newline bulk resistance & No & I \\

@@

            $<$ 2003 & Optimization for \newline cavity effects & reduces self-interference of quantum well & No & I, \cite{Shen2003} \\

@@

            ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\

@@

            1994 & double heterostructure & Higher radiative recombination probability & No & \cite{Nakamura1994} \\

@@

            1996 & multiple quantum well & Higher radiative recombination probability & No & \cite{Koike1996} \\

@@

            ongoing & Active region \newline Doping & radiative recombination probability & No & I, \cite{schubert2018light} \\

@@

            ongoing & Epitaxy \newline Improvements & radiative recombination probability & No & I \\

@@

            ongoing & Chip architecture \newline Improvements & radiative recombination probability & No & I \\

@@

            ongoing & Lower current density & Current density & No & I \\

@@

            ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\

@@

            ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\

@@

    \caption{LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article. \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.}

>>>>

The manufacturing process of semiconductor devices can be categorized by the level at which manufacturing process steps are implemented, i.e., either at the wafer level or at the individual chip/package level. The total manufacturing cost per die is thus the sum of the total costs of all wafer processing steps and all die packaging steps.
\begin{myequation}
\label{eqn:cost_sum}
    C \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = P_S + C_w + C_p
\end{myequation}
where
\begin{align*}
    P_S &\dots \text{sapphire substrate price per die} \\
    C_w &\dots \text{wafer processing cost per die} \\
    C_p &\dots \text{die processing cost}
\end{align*}
The total wafer processing cost and total die packaging costs are in turn the sum of all associated process steps.
\begin{myequation}
        C_w = \sum_i C_i
\end{myequation}
\begin{myequation}
	C_p = \sum_j C_j
\end{myequation}
The cost of a single process step $C_i$ can now be written as
\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{DPW}  \frac{1}{y_i}   \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x c_x \bigg\}
\end{myequation}
where the index $i$ runs over all wafer processing steps, the index $j$ runs over all die processing steps and the index $x$ run over all materials.
\begin{align*}
        DPW &\dots \text{number of functional (i.e., successfully tested) die per wafer}\\
        y &\dots \text{process step yield} \\
        u &\dots \text{equipment utilization (relative to theoretical equipment capacity)} \\
        p &\dots \text{power consumption} \\
        c_e &\dots \text{hourly electricity cost} \\
        c_m &\dots \text{hourly maintenance cost} \\
        c_d &\dots \text{hourly depreciation cost}\\
        c_l &\dots \text{hourly labour cost} \\
        c_o &\dots \text{hourly overhead cost} \\
        t_i &\dots \text{time per run} \\
        w &\dots \text{wafers per run} \\
        A &\dots \text{wafer area} \\
        v_x &\dots \text{volume of material $x$ per wafer} \\
        c_x &\dots \text{cost of material $x$ per volume}\\
\end{align*}
The number of die per wafer $D$ depends on the total usable wafer area. The usable area depends on the wafer diameter, the cutting street width between the chips and the exclusion zone at the rim of the wafer.
\begin{myequation}
	A_{\text{usable}}=A_{\text{wafer}}-A_{\text{cut}}-A_{\text{exclusion}}
\end{myequation}
Determining the usable wafer area as a function of these three parameters requires a numerical solution. However, following discussions in the literature \cite{de2005investigation}, we approximate the number of functional die per wafer\footnote{often \textit{"good die per wafer"} in the literature} as
\begin{myequation}
\label{eqn:dpw}
	DPW = \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2}
\end{myequation}
where
\begin{align*}
    d &\dots \text{wafer diameter} \\
    e &\dots \text{wafer edge exclusion zone width} \\
    a &\dots \text{die area} \\
    s &\dots \text{cutting street width} \\
\end{align*}

@@

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW) over time. \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category
\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x c_x \bigg\}
\end{split}
\end{myequation}
and the cost of a manufacturing step $C_j$ in the packaging category
\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x c_x \bigg\}
\end{myequation}
where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{DPW} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}
Note that in keeping with the categorization introduced by the United States Department of Energy (cf. eg. \cite{doe_ssl_rnd_2019}), certain steps from these two categories are reported separately. In the wafer processing category, the epitaxy step is reported separately due to its complexity and the large share of cost carried. In the wafer processing category, the phosphor step is reported separately.

@@

    \caption{Banana}

@@

            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\

@@

            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\

@@

            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\

@@

            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\

@@

            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\

@@

            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\

@@

            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\

@@

            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\

@@

            Ongoing & Lower current density & Current density & No & I \\

@@

            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\

@@

            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\

@@

    LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article. \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.
